The effects of binarity on planet occurrence rates measured by transit surveys

### 2017-12-12 sandbox/derivation.tex
improve precision
derivation: correct math errors

In a shell of thickness ${\rm d}r$, there are $n_d(r) 4\pi r^2 {\rm d}r$ 
searchable double star systems.
The number density of searchable double star systems $n_d(r)$ is constant up 
to the maximum searchable distance for singles, $d_0$. At greater distances, 
$n_d(r)$ decreases, because only some binaries have sufficient flux to make 
the SNR cut.
Specifically,
\begin{equation}
n_d(r) = 
\begin{cases}
n_b,	\quad r\leq d_0\\
n_b \int_{q_{\rm min}(r)}^1 {\rm d}q\, f_{\rm vl}(q), \quad d_0 < r < 
\sqrt{2}d_0 \\
\end{cases}
\end{equation}
where $f_{\rm vl}(q)$ is the mass ratio distribution in a volume-limited 
sample, $q_{\rm min}(r)$ is the minimum mass ratio required to provide 
sufficient flux for searchability.
Assuming $L\propto M^\alpha$,
\begin{equation}
q_{\rm min}(r) =
\left(
    \left( \frac{r}{d_0}\right)^2
    - 1
\right)^{\frac{1}{\alpha}}.
\end{equation}
Using the common parametrization $f_v(q) \sim q^\beta$, and setting $\beta=0$,
\begin{equation}
n_d(r) = 
\begin{cases}
n_b,	\quad r\leq d_0\\
n_b (1-q_{\rm min}(r)), \quad d_0 < r < \sqrt{2}d_0. \\
\end{cases}
\end{equation}
Integrating,
\begin{align}
N_d = \int {\rm d} N_d &= \int n_d(r) 4\pi r^2 {\rm d}r \\
&=
4\pi n_b \left(
    \frac{d_0^3}{3} +
    \left[\frac{r^3}{3}\right]^{\sqrt{2 d_0}}_{d_0}
    -
    \int_{d_0}^{\sqrt{2}d_0} \left(
    \left(\frac{r}{d_0}\right)^2 -1
    \right)^{\frac{1}{\alpha}}
    r^2 \,{\rm d}r
\right) \\
&=
\frac{4 \pi n_b}{3} d_0^3 \left(
2^{3/2} -
3 \int_{1}^{\sqrt{2}} {\rm d}u\, u^2 (u^2 - 1)^{1/\alpha}
\right), \label{eq:dimensionless_integral}
\\
&\equiv
\frac{4 \pi n_b}{3} d_0^3 \left(
2^{3/2} - 3 I
\right),
\end{align}
where the penultimate line substituted $u=r/d_0$.
For the common case $\alpha=3.5$, the dimensionless integral $I$ evaluates to 
$\approx 0.484174$.

@@

Power law planets are a slightly more realistic mass ratio distribution.
If we now assume
\begin{equation}
    \Gamma_i(r,M) = Z_i \times \frac{r^\delta}{\mathcal{N}_r} \times
    \frac{M^\gamma}{\mathcal{N}_M},
\end{equation}
where $Z_i$ is dimensionless, and the normalization constants carry 
the units (each side has units $[r^{-1} M^{-1}]$).
From Eq.~\ref{eq:simple_apparent_occ_rate}, we can write
\begin{equation}
\Gamma\a(r\a,M\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a,M\a) + \frac{{\rm BF}}{1-{\rm BF}} 
\left[ I_1(r_a,M\a) + I_2(r_a,M\a)
\right]
\right),
\end{equation}
where $I_1(r\a,M\a)$ and $I_2(r\a,M\a)$ are defined in the same way as in 
Eqs.~\ref{eq:defn_I1} and~\ref{eq:defn_I2}.
Doing the calculation, we find
\begin{align}
I_1(r\a,M\a) &= 
\int \frac{{\rm d}q}{\mathcal{A}} f(q)
\Gamma^1 \left(\frac{r\a}{\mathcal{A}},M\a\right)
\\
&=
\frac{Z_1}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+1}{2}}
\end{align}
and
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right)
\\
&= \frac{Z_2}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
               (1+q^\alpha)^{\frac{3}{2}}
               (1+q^{-\alpha})^{\frac{\delta+1}{2}}.
\end{align}
The net effect of binarity is to just {\it scale} the overall distribution!

Note that the true rate density for a volume-limited sample is
\begin{align}
\Gamma_{\rm vl}(r,M) &= \frac{r^\delta M^\gamma}{\mathcal{N}_r\mathcal{N}_M} 
\left[
\frac{Z_0}{1+2(n_b/n_s)} + \frac{(n_b/n_s)}{1+2(n_b/n_s)} (Z_1+Z_2)
\right] \\
&= \frac{r^\delta M^\gamma Z_0}{\mathcal{N}_r\mathcal{N}_M}
\quad\quad ({\rm assuming}\ Z_0=Z_1=Z_2),
\end{align}
which when $Z_0=Z_1=Z_2$ is equal to $\Gamma_0$, and $\Gamma_{\rm sel}$.
The correction factor with respect to single stars, $X_\Gamma = 
\Gamma\a/\Gamma_0$, is then
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a,M\a)}{\Gamma_0(r,M)} 
\right|_{r\a\rightarrow r,\ M\a\rightarrow M,\ Z_i{\rm 's\ equal}} \\
&=
\frac{1}{1+\mu}
\left[1 + \frac{2}{\mathcal{N}_q} \frac{{\rm BF}}{1-{\rm BF}}
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+1}{2}} +
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
    (1+q^\alpha)^{\frac{3}{2}}
    (1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\gamma=0$, $\delta=-2.92$, the sum of the 
integrals in $(\ldots)\approx 1.16393$. %171211_model_integrals.nb
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_0 = 1.36$, an 
apparent rate which is {\it higher} than the true rate by 36\%!
Note that this result is the same as $\Gamma\a/\Gamma_{\rm sel}$ and 
$\Gamma\a/\Gamma_{\rm vl}$ in the limit where all the $Z_i$'s are equal.

>>>>

In a shell of thickness ${\rm d}r$, there are $n_d(r) 4\pi r^2 {\rm d}r$ 
searchable double star systems.
The number density of searchable double star systems $n_d(r)$ is constant up 
to the maximum searchable distance for singles, $d_0$. At greater distances, 
$n_d(r)$ decreases, because only some binaries have sufficient flux to make 
the SNR cut.
Specifically,
\begin{equation}
n_d(r) = 
\begin{cases}
n_b,	\quad r\leq d_0\\
n_b \int_{q_{\rm min}(r)}^1 {\rm d}q\, f_{\rm vl}(q), \quad d_0 < r < 
\sqrt{2}d_0 \\
\end{cases}
\end{equation}
where $f_{\rm vl}(q)$ is the mass ratio distribution in a volume-limited 
sample, $q_{\rm min}(r)$ is the minimum mass ratio required to provide 
sufficient flux for searchability.
Assuming $L\propto M^\alpha$,
\begin{equation}
q_{\rm min}(r) =
\left(
    \left( \frac{r}{d_0}\right)^2
    - 1
\right)^{\frac{1}{\alpha}}.
\end{equation}
Using the common parametrization $f_v(q) \sim q^\beta$, and setting $\beta=0$,
\begin{equation}
n_d(r) = 
\begin{cases}
n_b,	\quad r\leq d_0\\
n_b (1-q_{\rm min}(r)), \quad d_0 < r < \sqrt{2}d_0. \\
\end{cases}
\end{equation}
Integrating,
\begin{align}
N_d = \int {\rm d} N_d &= \int n_d(r) 4\pi r^2 {\rm d}r \\
&=
4\pi n_b \left(
    \frac{d_0^3}{3} +
    \left[\frac{r^3}{3}\right]^{\sqrt{2 d_0}}_{d_0}
    -
    \int_{d_0}^{\sqrt{2}d_0} \left(
    \left(\frac{r}{d_0}\right)^2 -1
    \right)^{\frac{1}{\alpha}}
    r^2 \,{\rm d}r
\right) \\
&=
\frac{4 \pi n_b}{3} d_0^3 \left(
2^{3/2} -
3 \int_{1}^{\sqrt{2}} {\rm d}u\, u^2 (u^2 - 1)^{1/\alpha}
\right), \label{eq:dimensionless_integral}
\\
&\equiv
\frac{4 \pi n_b}{3} d_0^3 \left(
2^{3/2} - 3 I
\right),
\end{align}
where the penultimate line substituted $u=r/d_0$.
For the common case $\alpha=3.5$, the dimensionless integral $I$ evaluates to 
$\approx 0.4841741$.

@@

Power law planets are a slightly more realistic mass ratio distribution.
If we now assume
\begin{equation}
    \Gamma_i(r,M) = Z_i \times \frac{r^\delta}{\mathcal{N}_r} \times
    \frac{M^\gamma}{\mathcal{N}_M},
\end{equation}
where $Z_i$ is dimensionless, and the normalization constants carry 
the units (each side has units $[r^{-1} M^{-1}]$).
From Eq.~\ref{eq:simple_apparent_occ_rate}, we can write
\begin{equation}
\Gamma\a(r\a,M\a) = \frac{1}{1+\mu} \left(
\Gamma_0(r\a,M\a) + \frac{{\rm BF}}{1-{\rm BF}} 
\left[ I_1(r_a,M\a) + I_2(r_a,M\a)
\right]
\right),
\end{equation}
where $I_1(r\a,M\a)$ and $I_2(r\a,M\a)$ are defined in the same way as in 
Eqs.~\ref{eq:defn_I1} and~\ref{eq:defn_I2}.
Doing the calculation, we find
\begin{align}
I_1(r\a,M\a) &= 
\int \frac{{\rm d}q}{\mathcal{A}^4} f(q)
\Gamma^1 \left(\frac{r\a}{\mathcal{A}},M\a\right)
\\
&=
\frac{Z_1}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}}
\end{align}
and
\begin{align}
I_2(r\a,M\a) &=
\int \frac{q\,{\rm d}q}{\mathcal{A}^3 \mathcal{B}} f(q)
\Gamma^2 \left(\frac{r\a}{\mathcal{B}},qM\a\right)
\\
&= \frac{Z_2}{\mathcal{N}_q \mathcal{N}_r \mathcal{N}_M} 
r\a^\delta M\a^\gamma
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
               (1+q^\alpha)^{\frac{3}{2}}
               (1+q^{-\alpha})^{\frac{\delta+1}{2}}.
\end{align}
The net effect of binarity is to just {\it scale} the overall distribution!

Note that the true rate density for a volume-limited sample is
\begin{align}
\Gamma_{\rm vl}(r,M) &= \frac{r^\delta M^\gamma}{\mathcal{N}_r\mathcal{N}_M} 
\left[
\frac{Z_0}{1+2(n_b/n_s)} + \frac{(n_b/n_s)}{1+2(n_b/n_s)} (Z_1+Z_2)
\right] \\
&= \frac{r^\delta M^\gamma Z_0}{\mathcal{N}_r\mathcal{N}_M}
\quad\quad ({\rm assuming}\ Z_0=Z_1=Z_2),
\end{align}
which when $Z_0=Z_1=Z_2$ is equal to $\Gamma_0$, and $\Gamma_{\rm sel}$.
The correction factor with respect to single stars, $X_\Gamma = 
\Gamma\a/\Gamma_0$, is then
\begin{align}
X_\Gamma &\equiv \left. \frac{\Gamma\a(r\a,M\a)}{\Gamma_0(r,M)} 
\right|_{r\a\rightarrow r,\ M\a\rightarrow M,\ Z_i{\rm 's\ equal}} \\
&=
\frac{1}{1+\mu}
\left[1 + \frac{1}{\mathcal{N}_q} \frac{{\rm BF}}{1-{\rm BF}}
\left(
\int {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\delta+4}{2}} +
\int {\rm d}q\,q^{\beta+\gamma+\delta+\frac{8}{3}} 
    (1+q^\alpha)^{\frac{3}{2}}
    (1+q^{-\alpha})^{\frac{\delta+1}{2}}
\right)\right].
\end{align}

For $\alpha = 3.5$, $\beta=0$, $\gamma=0$, $\delta=-2.92$, the sum of the 
integrals in $(\ldots)\approx 1.41425$. %171211_model_integrals.nb
For ${\rm BF}=0.44$, this yields $\Gamma\a/\Gamma_0 = 1.01448$, an 
apparent rate which is {\it higher} than the true rate by 1.4\%!
Note that this result is the same as $\Gamma\a/\Gamma_{\rm sel}$ and 
$\Gamma\a/\Gamma_{\rm vl}$ in the limit where all the $Z_i$'s are equal.
